### 2016-04-20 paper/top_words_table.tex
Added top words table.

\setlength{\tabcolsep}{4pt}
\begin{tabular}{cccccccccc}
  \toprule
  Topic 1 & Topic 2 & Topic 3 & Topic 4 & Topic 5   & Topic 6     & Topic 7  & Topic 8 & Topic 9   & Topic 10 \\
  \midrule
  call    & mistr   & know    & see     & state     & israel      & peopl    & think   & meet      & discuss  \\
  huma    & last    & just    & report  & benghazi  & isra        & world    & like    & reuters   & letter   \\
  abedin  & said    & want    & haiti   & inform    & peac        & power    & good    & follow    & draft    \\
  schedul & parti   & let     & news    & date      & palestinian & polit    & say     & update    & cdm      \\
  sheet   & david   & ask     & great   & subject   & netanyahu   & mani     & dont    & list      & note     \\
  request & prime   & come    & read    & agreement & east        & right    & look    & happi     & eam      \\
  speak   & deal    & tri     & media   & depart    & negoti      & can      & got     & monday    & secretar \\
  readout & leader  & sure    & event   & doc       & arab        & govern   & tell    & scshedule & question \\
  calls   & govern  & thx     & post    & hous      & state       & one      & thing   & set       & final    \\
  phone   & ireland & best    & connect & produc    & middl       & american & idea    & mtg       & brief    \\
  \bottomrule \\
  \toprule
  Topic 11   & Topic 12  & Topic 13 & Topic 14 & Topic 15 & Topic 16   & Topic 17   & Topic 18 & Topic 19 & Topic 20 \\
  \midrule
  democrat   & women     & new      & will     & make     & secretari  & senat      & libya    & iran     & time \\
  republican & clinton   & foreign  & work     & well     & depart     & bill       & secur    & syria    & press \\
  american   & right     & trip     & week     & take     & office     & vote       & travel   & nuclear  & hrc \\
  polit      & hillari   & told     & need     & way      & room       & health     & libyan   & syrian   & staff \\
  parti      & secretari & import   & next     & point    & state      & care       & iraq     & turkey   & dinner \\
  elect      & year      & agre     & also     & hope     & arriv      & pass       & embassi  & russia   & lona \\
  obama      & state     & high     & plan     & time     & rout       & hous       & attack   & regim    & ambassador \\
  percent    & human     & head     & start    & move     & privat     & john       & kill     & iranian  & close \\
  candid     & day       & later    & issu     & possibl  & offic      & committe   & march    & intern   & monica \\
  poll       & mani      & spoke    & keep     & one      & department & republican & august   & opposit  & remark \\
  \bottomrule \\
  \toprule
  Topic 21 & Topic 22 & Topic 23  & Topic 24 & Topic 25  & Topic 26 & Topic 27    & Topic 28 & Topic 29  & Topic 30 \\
  \midrule
  fyi      & said     & develop   & email    & presid    & china    & afghanistan & can      & statement & tomorrow \\
  print    & book     & state     & offic    & obama     & econom   & pakistan    & talk     & egypt     & today \\
  sid      & one      & support   & sent     & said      & year     & afghan      & get      & egyptian  & cheryl \\
  thank    & year     & global    & messag   & hous      & chines   & militari    & speech   & respond   & sullivan \\
  memo     & two      & program   & pleas    & white     & job      & general     & now      & announc   & jacob \\
  pls      & day      & effort    & may      & administr & million  & karzai      & back     & releas    & wil \\
  latest   & case     & intern    & check    & polici    & countri  & offici      & updat    & elect     & morn \\
  info     & famili   & work      & copi     & aid       & bank     & war         & send     & saudi     & jake \\
  intel    & school   & includ    & forward  & advis     & polici   & forc        & tonight  & reuter    & mill \\
  urgent   & court    & diplomaci & blair    & former    & economi  & nato        & soon     & presid    & confirm \\
  \bottomrule
\end{tabular}
\setlength{\tabcolsep}{6pt}